Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-12-18 xrb2/paper.tex
added to section 3.2

The mean molecular weight $\bar{A}$ within each of our simulations seems to grow along defined tracks confined to certain temperatures $T$, as shown in the Figure \ref{fig:abar} phase plots. We believe that the tracks in the plot correspond to reaction pathways in the reaction network. Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined. The ''messiness" of the tracks may be dependent on how mixed the flame interior is --- since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}. {\tt F1000} aligns with this interpretation: its $\bar{A}$ tracks are somewhat disrupted compared to the slower rotation runs, possibly due to the more vigorous mixing of the vortex at the flame front. Comparing the different runs, we also see that as the rotation rate increases, so does the peak temperature. This makes sense if higher rotation leads to a more confined, intense vortex near the flame front. It also agrees with our earlier interpretation of enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}. The disruption of reaction pathways with higher rotation rates and mixed flames may have interesting consequences; if the reactions are disrupted enough, it is possible that the burning could be snuffed out. This can occur in combustion scenarios when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt the burning.

@@

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate. As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters. These results are given in Table~\ref{table:flame_speeds}. The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$. \MarginPar{How is the acceleration calculated for this table? I.e. Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to four
to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.

>>>>

The mean molecular weight $\bar{A}$ within each of our simulations seems to grow along defined tracks confined to certain temperatures $T$, as shown in the Figure \ref{fig:abar} phase plots. We believe that the tracks in the plot correspond to reaction pathways in the reaction network. Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined. The ''messiness" of the tracks may be dependent on how mixed the flame interior is --- since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}. {\tt F1000} aligns with this interpretation: its $\bar{A}$ tracks are somewhat disrupted compared to the slower rotation runs, possibly due to the more vigorous mixing of the vortex at the flame front. Comparing the different runs, we also see that as the rotation rate increases, so does the peak temperature. This makes sense if higher rotation leads to a more concentrated, intense vortex near the flame front. It also agrees with our earlier interpretation of enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}. The disruption of reaction pathways with higher rotation rates and mixed flames may have interesting consequences; if the reactions are disrupted enough, it is possible that the burning could be snuffed out. This can occur in combustion scenarios when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt the burning.

@@

Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs (blue and orange, respectively) to show the dependence on rotation rate. As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for times $t \gtrsim 0.02~\mathrm{s}$ and calculate the acceleration and initial velocity from the fit parameters. These results are given in Table~\ref{table:flame_speeds}. \MarginPar{How is the acceleration calculated for this table? I.e. Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.

We found in \citet{flame_wave1} that the flame speed $s$ obeys:
\begin{equation}
	\label{eqn:rossby}
	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
\end{equation}
where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the increased Coriolis confinement. The initial position of $\tt{F1000}$ does appear to be reduced compared to the slower $\tt{F500}$. From Equation \ref{eqn:rossby}, increasing the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial position of the flame should be reduced by a similar factor. Qualitatively, we think that the higher rotation $\tt{F1000}$ run starts with a tighter votex due to the increased Coriolis force, leading to a lower initial position for the flame front. $L_R$ is not, however, a perfect metric for the flame lengthscale. The scaling described in Equation \ref{eqn:rossby} seems to be more accurate at higher rotation rates (1000 Hz and up), based on \citet{flame_wave1}. We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
